---
title: "Bayesian Analysis of Vertical Words and Letters experiment, For Preregistration"
author: "Alex Holcombe"
date: '`r paste("Mostly done by Aug 15, 2019. Updated on", Sys.Date())`'

output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary

This is for an experiment by Kim Ransley and Alex Holcombe in which in one condition, two letters are presented in vertical configuration, and in other condition, two words are presented in vertical configuration (half of trials in left visual field and half in right visual field, but will probably collapse across the two visual fields for most of these analyses).

The prereg of the first attempt at the experiment with the incorrect (due to programming error) vertical alignment is bayesianAnalysisVerticalForPrereg.Rmd.




## Prediction: No diff between upper and lower words for position of errors.

Analysis of the spatial position of errors on the top and bottom. Our interest was whether the average letter position of errors was closer to the end of the word for responses in the bottom stream than the top stream. This could indicate that there was a letter-by-letter scanning process rather than the processing occurring at the whole-word level, because with a letter-by-letter account, there might not be time to get to the end of the lower word. So our prediction is that the data will favor the null hypothesis 

We will calculate the average letter position of errors for each trial (for example, trap rather than tram is a position 4 error) and averaged these for each participant.  We did this only for those trials where participants had made responses that were orthographically similar to the target (as measured by a score of 0.7 [MAYBE WE SHOULD SET THIS HIGHER?] on the van Orden (1987) orthographic similarity index, which generally accepts answers where participants get two or more of the letters correct). We then used a one-tailed Bayesian t-test to test whether the average letter position of errors was higher for responses in the right stream than the left stream. 


### Dependent variables



### Considerations for the prior on the effect size

In E1 (horizontally arrayed letters versus horizontally arrayed words), the left advantage of the letters was smaller than in all previous experiments, which is a bit worrying. It could be a fluke, or it could be because of half of Ss just done the words which givs a big right hemifield advantage, or it could be the oral response. All additional reasons to be conservative.

Another difference with our previous work is that in this experiment both streams are in one or the other hemisphere.

## Bayesian paired t-test with custom prior

```{r, echo = FALSE}
library(BayesFactor) #Morey's package


```