---
title: "MBL day 1"
author: "Alex Holcombe"
date: "2/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library('mixRSVP')
```

Analysis of MBL day 1.

Script based mostly on https://github.com/alexholcombe/Cheryl_E2/blob/master/dataPreprocess/readInE2rawData.R

```{r set up functions}

rawDataPath<- file.path("..","data","MBL","firstDay")

files <- dir(path=rawDataPath,pattern='.txt')  #find all data files in this directory
dfAll<-data.frame()
print(files)
```


```{r all}

for (i in 1:length(files)) { #read in each file
    fileThis<- file.path(rawDataPath,files[i])

    rawData<- tryCatch(
       read.table( fileThis,  sep="\t", header=TRUE, strip.white=TRUE),
       error=function(e) {
         stop( paste0("ERROR reading the file ",fname," :",e) )
       } )

    filename<-files[i]
    apparentSubjectName <- strsplit(filename,split="_")[[1]][1]
    print(paste("Apparent subject name about to load is",apparentSubjectName))
    subjectName<- as.character(  rawData$subject[1] )
    if (apparentSubjectName != subjectName) {
      stop( paste("WARNING apparentSubjectName",apparentSubjectName," from filename does not match subjectName in data structure",
                  subjectName, "in file",files[i]) )
    }
    rawData$file <- filename

    dfThis<-rawData # turnRawPsychopyOutputIntoMeltedDataframe(rawData)

    tryCatch(
      dfAll<-rbind(dfAll,dfThis), #if fail to bind new with old,
      error=function(e) { #Give feedback about how the error happened
        cat(paste0("Tried to merge but error:",e) )
      } )
      
}
```

Add SOA and letterDur to data, because I tested MBL at multiple values

```{r hey}
library(dplyr)

dfAll<- dfAll %>% mutate( SOA = case_when(file=="MBL_26Apr2019_11-43.txt" ~ 360,
                                          file=="MBL_26Apr2019_11-50.txt" ~ 360,
                                          file=="MBL_26Apr2019_11-54.txt" ~ 400,
                                          file=="MBL_26Apr2019_11-56.txt" ~ 600,
                                          file=="MBL_26Apr2019_11-58.txt" ~ 900,
                                          file=="mbl_26Apr2019_12-01.txt" ~ 350 #almost all the data is in this file
                                          ))

dfAll<- dfAll %>% mutate( letterdurDesired = case_when(file=="MBL_26Apr2019_11-43.txt" ~ 80,
                                                file=="MBL_26Apr2019_11-50.txt" ~ 80,
                                                file=="MBL_26Apr2019_11-54.txt" ~ 80,
                                                file=="MBL_26Apr2019_11-56.txt" ~ 80,
                                                file=="MBL_26Apr2019_11-58.txt" ~ 80,
                                                file=="mbl_26Apr2019_12-01.txt" ~ 300) )


refreshRate <- 60
dfAll<- dfAll %>% mutate(   letterDurFrames = floor (letterdurDesired / (1000./refreshRate) )  )
#table(dfAll$letterdur, dfAll$letterDurFrames)
dfAll$letterdur <- dfAll$letterDurFrames * 1000/refreshRate

table(dfAll$SOA, dfAll$letterdur)                 
```

```{r visual all rates check}
library(ggplot2)


sanityVisualCheck<-TRUE
if (sanityVisualCheck) {

  g=ggplot(dfAll, aes(x=responsePosRelative0))
  #plot data
  g<-g+geom_histogram(binwidth=1) + xlab('serial position error') + facet_grid(.~SOA) + ggtitle('left letter')
  g <- g + geom_text(aes(x=-10,y=0,label=round(letterdur,0))) #paste("dur:",)))
  g<-g+ xlim(-11,11) #18 items per stream. Based on Kim's program so included letters that were never presented in the stream
  show(g)
  
  h=ggplot(dfAll, aes(x=responsePosRelative1))
  #plot data
  h<-h+geom_histogram(binwidth=1) + xlab('serial position error') +  facet_grid(.~SOA) + ggtitle('right letter')
  h <-h + geom_text(aes(x=-10,y=0,label=round(letterdur,0))) #paste("dur:",)))
  h<-h+ xlim(-11,11) #18 items per stream. Based on Kim's program so included letters that were never presented in the stream
  show(h)
  
}
```

Only analyse at 350 ms SOA because for all others letterdur was 80 ms and he couldn't do it.

```{r visual check}

E<- dfAll %>% filter(SOA==350, experimentPhase=="main")

sanityVisualCheck<-TRUE
if (sanityVisualCheck) {
  library(ggplot2)

  g=ggplot(E, aes(x=responsePosRelative0))
  #plot data
  g<-g+geom_histogram(binwidth=1) + xlab('serial position error') + ggtitle('left letter')
  g<-g+ xlim(-11,11) #18 items per stream. Based on Kim's program so included letters that were never presented in the stream
  show(g)
  
  g=ggplot(E, aes(x=responsePosRelative1))
  #plot data
  g<-g+geom_histogram(binwidth=1) + xlab('serial position error') + ggtitle('right letter')
  g<-g+ xlim(-11,11) #18 items per stream. Based on Kim's program so included letters that were never presented in the stream
  show(g)
}
```


Melt.  Error "attributes are not identical across measure variables; they will be dropped" seems OK.

```{r melt with me}
library(tidyr)

columnsToMelt<- colnames(E)[11:20]
#https://stackoverflow.com/questions/23945350/reshaping-wide-to-long-with-multiple-values-columns
dl<-  E %>%  gather(v, value, columnsToMelt)
#attributes dropped but that's ok
#Separate the new v column into 0 and 1 (corresponding to left and right)
dl<- dl %>% separate(col = v, into = c("col", "var"), sep = -1) #This means separate it at the last character, 
  #which is always 0 or 1 corresponding to left and right
dl<- dl %>% arrange(file,trialnum) %>% 
  spread(col, value) %>% rename(side=var, SPE=responsePosRelative)
rownames(dl)<-NULL

dl$SPE <- as.numeric(dl$SPE)


```

```{r calc needed experiment params}
numItemsInStream <- 17
dl$cueSerialPos <- as.numeric(dl$cueSerialPos) 
possibleTargetSP<- sort(unique(dl$cueSerialPos))
minTargetSP <- min( possibleTargetSP ) 
maxTargetSP <- max( possibleTargetSP )
minSPE <- 1 - maxTargetSP
maxSPE <- numItemsInStream - minTargetSP
```


Plot after melting
```{r}
dl[dl$side=="0",]$side <- "left"
dl[dl$side=="1",]$side <- "right"

g=ggplot(dl, aes(x=SPE)) + theme_apa() + facet_grid(side~.)
#plot data
g<-g+geom_histogram(binwidth=1) + xlim(minSPE,maxSPE)
show(g)
```

```{r}
dl<- dl %>% rename(targetSP = cueSerialPos)
#WHY DOES THE targetSP variable have to be called targetSP for plot_hist_with_fit when you have to pass the motherfucker?
#Showit also doesn't seem to be working
```

Prepare for mixture modeling by putting the few -999 errors (where he reported a letter not in the stream) as a random error not close to the peak. This is a conservative kludge. Otherwise have to modify the mixture modeling code to model -999 or another dummy value as pure guessing, which is doable but probably would be v. time-consuming. Don't have time now, better to wait until Mike writing this up.

```{r}
guessesFakeSPEs <- minSPE:maxSPE
guessesFakeSPEs <- guessesFakeSPEs[-pmatch(-3:3,guessesFakeSPEs)] #remove -3:3 to be sure they're modeled as part of the uniform distribution #https://stackoverflow.com/questions/49742315/how-to-delete-multiple-values-from-a-vector-but-not-all-of-them
dl[dl$SPE==-999,]$SPE <- sample(guessesFakeSPEs, size=sum(dl$SPE==-999),replace=TRUE)

```

Mixture model the left side

```{r mixml}
dlLeft<- dl %>% filter(side=="left") 

plotContinuousGaussian <- FALSE; annotateIt <- TRUE
j<- plot_hist_with_fit(dlLeft,minSPE,maxSPE,dlLeft$targetSP,numItemsInStream,plotContinuousGaussian,annotateIt, FALSE)
j<- j + ggtitle("left")
show(j)
```

Mixture model the right side
```{r}
dlRight<- dl %>% filter(side=="right") %>% filter(SPE>-999)  #ignore guesses of letters that weren't in the stream. Could alternatively give them random values

plotContinuousGaussian <- TRUE; annotateIt <- TRUE
k<- plot_hist_with_fit(dlRight,minSPE,maxSPE,dlRight$targetSP,numItemsInStream,plotContinuousGaussian,annotateIt, FALSE)
k<- k + ggtitle("right")
show(k)
```

Save the data, for combining later with day two.
```{r save file}
library(readr) #has better than save.csv https://readr.tidyverse.org/reference/write_delim.html
pathToSave<- file.path("..","data","MBL")

saveRDS(dl, file = file.path(pathToSave,"dayOne.rda"), compress="bzip2")
write_tsv(dl, file.path(pathToSave,"dayOne.tsv"))
```
